#Johns Hopkins University- Data Science Specialization

This is the first project in Reproducible Research in *Johns Hopkins University- Data Science Specialization*.The dataset in this project is downloaded from the github of the instructor, Roger D. Peng, readers can also download the dataset from [my github page](https://github.com/CaesarLWY/RepData_PeerAssessment1) where I had forked Roger's.  

The dataset is downloaded on 2018-02-28 23:01.

##1.Loading and preprocessing the data
###Load the data 

```{r read}
setwd("C:/Users/c1992/Desktop/MOOCS-20170607T120517Z-001/MOOCS/JHU- Data Science/5- Reproducible Research")
data <- read.csv("./2_project_data_activity/activity.csv")
head(data)
dim(data)
```
There are 17568 rows(observations), but with only 3 variables.
Check the variable steps to count how many NAs in the data:

```{r check}
sum(is.na(data$steps))
```

Apparently, there are 2304 NAs in the dataset, the observations with NA in steps variable should be eliminated for the purpose of further analysis.

###Process the data into a format suitable for the analysis
```{r tidfy}
tidydata <- data[-which(is.na(data$steps)),]
head(tidydata)
tail(tidydata)
sum(is.na(tidydata$steps))
```

##2.What is mean total number of steps taken per day?

###Calculate the total number of steps taken per day.

```{r plot}
data_daily <- aggregate(list(tidydata$steps), by= list(tidydata$date), FUN = sum)
names(data_daily) <- c("Date","counts")
head(data_daily)
```

###Make a histogram of the total number of steps taken each day.

```{r histogram}
hist(data_daily$counts)
```

###Calculate and report the mean and median of the total number of steps taken per day.

```{r mean&median}
data_daily_mean <- aggregate(list(tidydata$steps), by= list(tidydata$date), FUN= mean)
names(data_daily_mean) <- c("date","mean") 
data_daily_mean
data_daily_median <- aggregate(list(tidydata$steps), by= list(tidydata$date), FUN= median)
names(data_daily_median) <- c("date", "median")
data_daily_median

```

##3.What is the average daily activity pattern?

###Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
```{r time series}
avgdata <- aggregate(list(tidydata$steps), by= list(tidydata$interval), FUN= mean)
names(avgdata) <- c("time","avg.steps")
head(avgdata)
with(avgdata, plot(avg.steps~time, type="l"))
```

###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  
To find which time interval across all the days possess the maximum value, we can run the following code to achieve:

```{r max}
avgdata[which.max(avgdata$avg.steps),]
```
We can observe that when time interval is 835, avg.steps reaches its peak on 206.1698.
 
##4.Imputing missing values
###Calculate and report the total number of missing values in the dataset.
As the analysis in part1.Reading and Processing data, there are 2304 NAs in the preliminary dataset.Instea of ommiting these observations, we can use another way to deal with these observations with NAs.

###Devise a strategy for filling in all of the missing values in the dataset. 
```{r impute}
library(impute)
date <- data[,2]
data.imputed <- impute.knn(as.matrix(data[,-2]))$data
data.imputed <- as.data.frame(data.imputed)
sum(is.na(data.imputed$steps))
head(data.imputed)
tail(data.imputed)
```

In the previous code chunk, *k-nearest neighbor* is adopted to impute the missing values, and by the result of sum(is.na(data.imputed$steps))==0, we can easily observe the NAs in the steps variable had fallen from 2304 to 0.

###Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r add date}
data.imputed <- cbind(data.imputed, date)
head(data.imputed)
tail(data.imputed)
```

Now we have the complete data table being similar with the original one, except plugging into simulated values to replace NAs.

###Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.Do these values differ from the estimates from the first part of the project? What is the impact of imputing missing data on the estimates of the total daily number of steps? 
```{r hist daily steps}
data.imputed_count <- aggregate(list(data.imputed$steps), by= list(data.imputed$date), FUN= sum)
names(data.imputed_count) <- c("date", "counts")
head(data.imputed_count)
hist(data.imputed_count$counts, xlab= "Daily steps", main= "Histogram of daily steps")
```

Then calculate the mean value and median value of daily steps without NAs.

```{r mean&median.imputed}
data.imputed_mean <- aggregate(list(data.imputed$steps), by= list(data.imputed$date), FUN= mean)
names(data.imputed_mean) <- c("date","mean")
data.imputed_mean

data.imputed_median <- aggregate(list(data.imputed$steps), by= list(data.imputed$date), FUN= median)
names(data.imputed_median) <- c("date","median")
data.imputed_median
```

Compare the result of dataset omitting NAs with dataset imputing NAs. We draw the daily steps histogram out to compare here.

```{r compare hist}
par(mfrow= c(1,2))
hist(data_daily$counts, xlab= "Daily steps", main= "Omitting NAs")
hist(data.imputed_count$counts, xlab= "Daily steps", main= "Imputing NAs")
```
In the previous plots, we can observe that the frequency of daily steps increases from 12 to 20 in the interval(5000,10000).There are also some changes in median and mean of daily steps, including 0s in the median tables decreses,...etc.

##5.Are there differences in activity patterns between weekdays and weekends?

###Create a new factor variable in the dataset with two levels ¡V ¡§weekday¡¨ and ¡§weekend¡¨ indicating whether a given date is a weekday or weekend day.

```{r weekdays}
Sys.setlocale("LC_TIME","English")
data.imputed$weekdays <-
weekdays(as.POSIXct(data.imputed$date))
head(data.imputed)
tail(data.imputed)
```

```{r assign levels}
class(data.imputed$weekdays)
data.imputed$weekdays <- as.factor(data.imputed$weekdays)
levels(data.imputed$weekdays)
levels(data.imputed$weekdays)[levels(data.imputed$weekdays)==c("Saturday","Sunday")] <- "weekend"
levels(data.imputed$weekdays)
levels(data.imputed$weekdays)[levels(data.imputed$weekdays)!="weekend"] <- "weekday"
levels(data.imputed$weekdays)
```

###Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r time series 2}
library(ggplot2)
g <- ggplot(data.imputed, aes(interval, steps))
g + geom_line()+ facet_grid(.~weekdays) 
```




